---
title: "UN Datathon work"
author: "Meg Ellingwood"
date: "2023-11-03"
output: html_document
---

```{r}
library(tidyverse)
library(ggmap)

```
(Very informal messing around so far)
```{r}
spatial_confounding_data <- read_delim("~/Documents/extra-projects/un-datathon-2023/analysis_dat.tab")
```

Setting up a US map
```{r}
# US <- c(left = -125, bottom = 24,
#         right = -67, top = 49)
# 
# register_stadiamaps("682d676b-3e06-4933-ad87-65994afa1362")
# 
# US_map <- get_map(location = c(lon = -95.3632715, lat = 29.7632836), zoom = "auto",
#                      source = "osm")
# 
# ggmap(US_map)
```
ggmap doesn't work anymore oof

```{r}
library(tmap)

data("World")
tm_shape(World) +
  tm_polygons("HPI")

```
(From tmap vignette)
```{r}
data(World, rivers)

qtm(rivers)

## Not run: 
tm_shape(World) +
    tm_fill() +
tm_shape(rivers) +
    tm_lines(col="black", lwd="scalerank", scale=2, legend.lwd.show = FALSE) +
tm_style("cobalt", title = "Rivers of the World") +
tm_format("World")

## End(Not run)
```
So their native datasets might not be super informative but the framework could be useful


tutorial by https://milospopovic.net/crisp-topography-map-with-r/
Simply making a topographical map of a given country
```{r}
# libraries we need
libs <- c("elevatr", "terra", "tidyverse", 
	"sf", "giscoR", "marmap")

# install missing libraries
installed_libs <- libs %in% rownames(installed.packages())
if (any(installed_libs == F)) {
  install.packages(libs[!installed_libs])
}

# load libraries
invisible(lapply(libs, library, character.only = T))

crsLONGLAT <- "+proj=longlat +datum=WGS84 +no_defs"

get_sf <- function(country_name = "Italy", country_transformed, res = "10") {
	
	country_sf <- giscoR::gisco_get_countries(
    	year = "2016",
    	epsg = "4326",
    	resolution = res,
    	country = country_name)
	
	country_transformed <- st_transform(country_sf, crs = crsLONGLAT)

	return(country_transformed)
}

country_transformed <- get_sf()
```


```{r}
get_elevation_data <- function(country_transformed, country_elevation_df, res = 9) {

	country_elevation <- get_elev_raster(
		locations = country_transformed, 
		z = res, 
		clip = "locations") 

	country_elevation_df <- as.data.frame(country_elevation, xy = T) %>%
		na.omit()
	
	colnames(country_elevation_df)[3] <- "elevation"

	return(country_elevation_df)
}

country_elevation_df <- get_elevation_data(country_transformed)
```

```{r}
get_elevation_map <- function(country_elevation_df) {

	country_map <- ggplot() +
  		geom_tile(data = country_elevation_df, 
  			aes(x = x, y = y, fill = elevation)) +
  		scale_fill_etopo() +
	  coord_sf(crs = crsLONGLAT)+
  		theme_minimal() +
  		theme(text = element_text(color = "#22211d"),
    		axis.line = element_blank(),
    		axis.text.x = element_blank(),
    		axis.text.y = element_blank(),
    		axis.ticks = element_blank(),
    		axis.title.x = element_blank(),
    		axis.title.y = element_blank(),
    		legend.position = "none",
   		  	panel.grid.major = element_line(color = "white", size = 0.2),
    		panel.grid.minor = element_blank(),
    		plot.title = element_text(size=18, color="grey20", hjust=1, vjust=-5),
    		plot.caption = element_text(size=8, color="grey70", hjust=.15, vjust=20),
    		plot.margin = unit(c(t=0, r=0, b=0, l=0),"lines"), #added these narrower margins to enlarge maps
    		plot.background = element_rect(fill = "white", color = NA), 
    		panel.background = element_rect(fill = "white", color = NA),
    		panel.border = element_blank()) +
		labs(x = "", 
    		y = NULL, 
    		title = "Topographic map of ITALY")

	return(country_map)
}

country_map <- get_elevation_map(country_elevation_df)

country_map
```

Try another country?
```{r}
brazil_transformed <- get_sf("Brazil", res = "10")
brazil_elevation <- get_elevation_data(brazil_transformed, res = 4)
get_elevation_map(brazil_elevation)
```


Can I do this a bit more simply with just accessing the package functions directly?

Gotta get full rectangle data:
```{r}
library(rnaturalearth)

target_crs <- "+proj=longlat +datum=WGS84 +no_defs"

map <- ne_download(scale = "large", type = "countries",
                   returnclass = "sf")
map <- st_transform(map, crs = target_crs)
```




FOCUSING ON BRAZIL
Brazil lat-long coords: top = 5, bottom = -35, left = -75, right = -35
```{r}
brazil_latlong <- c(left = -75, bottom = -35,
       right = -35, top = 5)

brazil_window <- st_sfc(st_point(c(-75, -35)), #left, bottom
                        st_point(c(-35, 5)), #right, top
                        crs = 4326)

brazil_window_sf <- brazil_window %>% 
  st_transform(crs = target_crs) %>% 
  st_coordinates()

region_brazil <- c("Brazil", "French Guiana", "Suriname", "Guyana",
                   "Venezuela", "Colombia", "Peru", "Bolivia", 
                   "Paraguay", "Argentina", "Uruguay", "France")

map <- map %>% 
  filter(SOVEREIGNT %in% region_brazil)
```
So that gives me the sf dataframe with the outline of Brazil?
```{r}
ggplot(map) +
  geom_sf() +
  coord_sf(xlim = brazil_window_sf[, "X"],
           ylim = brazil_window_sf[, "Y"]) 
```


Yay for appropriate region map 
Now for the actual elevation data
```{r}
br_elevation <- get_elev_raster(
		locations = map, 
		z = 5, 
		clip = "locations") 

br_elevation_df <- as.data.frame(br_elevation, xy = T) %>%
		na.omit()

colnames(br_elevation_df)[3] <- "elevation"

br_elevation_df <- br_elevation_df %>% 
  filter(x <= -30, x >= -80, y <= 10, y >= -40) %>% 
  mutate(elevation = ifelse(elevation > 3000, 3000, elevation), 
         elevation = ifelse(elevation < 0, NA, elevation))

```
Note: cutting off elevation at 3000m because the highest point in Brazil is this mountain, so as not to get thrown off by the Andes to the west
https://en.wikipedia.org/wiki/Pico_da_Neblina#:~:text=Pico%20da%20Neblina%20(Portuguese%20pronunciation,on%20the%20Brazil%2DVenezuela%20border.



Mapping the elevation
```{r}
ggplot(map) +
  geom_sf(colour = "black") +
  coord_sf(xlim = brazil_window_sf[, "X"],
           ylim = brazil_window_sf[, "Y"]) +
  geom_tile(data = br_elevation_df, 
  			aes(x = x, y = y, fill = elevation)) +
  scale_fill_gradient(low = "white", high = "navy")
```
```{r}
write_csv(br_elevation_df, "~/Documents/extra-projects/un_datathon_2023/DATA/elevation_res5.csv")
```


Now making the boundary box smaller
```{r}
rio_window <- st_sfc(st_point(c(-47.5, -24)), #left, bottom
                        st_point(c(-42.5, -19)), #right, top
                        crs = 4326)

rio_window_sf <- rio_window %>% 
  st_transform(crs = target_crs) %>% 
  st_coordinates()

rio_elev_df <- br_elevation_df %>% 
  filter(y > -24 & y < -19 & x < -42.5 & x > -47.5)
```
Displaying
```{r}
plot_elev <- ggplot(map) +
  geom_tile(data = rio_elev_df, 
  			aes(x = x, y = y, fill = elevation)) +
  scale_fill_gradient(low = "white", high = "navy") +
  geom_sf(alpha = 0.1) +
  coord_sf(xlim = rio_window_sf[, "X"],
           ylim = rio_window_sf[, "Y"]) +
  labs(x = "Longitude", y = "Latitude", fill = "Elevation (m)") +
  theme_bw()
plot_elev
ggsave("~/Documents/extra-projects/un_datathon_2023/VIZ/elevation_plot.jpg")
```

