---
title: "UN Datathon work"
author: "Meg Ellingwood"
date: "2023-11-03"
output: html_document
---

```{r}
## library(ggmap) ## ggmap not currently working with stamenmap
library(tmap) ## didn't end up using here, but useful
library(elevatr)
library(terra)
library(tidyverse)
library(sf)
library(giscoR)
library(marmap)

```

Initially tried following tutorial by https://milospopovic.net/crisp-topography-map-with-r/
Simply making a topographical map of a given country


Then switched to rnaturalearth package to be able to get full rectangle around the country of interest

Get full rectangle data:
```{r}
library(rnaturalearth)

target_crs <- "+proj=longlat +datum=WGS84 +no_defs"

map <- ne_download(scale = "large", type = "countries",
                   returnclass = "sf")
map <- st_transform(map, crs = target_crs)
```


FOCUSING ON BRAZIL
Brazil lat-long coords: top = 5, bottom = -35, left = -75, right = -35
```{r}
brazil_latlong <- c(left = -75, bottom = -35,
       right = -35, top = 5)

brazil_window <- st_sfc(st_point(c(-75, -35)), #left, bottom
                        st_point(c(-35, 5)), #right, top
                        crs = 4326)

brazil_window_sf <- brazil_window %>% 
  st_transform(crs = target_crs) %>% 
  st_coordinates()

region_brazil <- c("Brazil", "French Guiana", "Suriname", "Guyana",
                   "Venezuela", "Colombia", "Peru", "Bolivia", 
                   "Paraguay", "Argentina", "Uruguay", "France")

map <- map %>% 
  filter(SOVEREIGNT %in% region_brazil)
```

So that gives me the sf dataframe with the outline of Brazil
```{r}
ggplot(map) +
  geom_sf() +
  coord_sf(xlim = brazil_window_sf[, "X"],
           ylim = brazil_window_sf[, "Y"]) 
```

Now for the actual elevation data
```{r}
br_elevation <- get_elev_raster(
		locations = map, 
		z = 5, 
		clip = "locations") 

br_elevation_df <- as.data.frame(br_elevation, xy = T) %>%
		na.omit()

colnames(br_elevation_df)[3] <- "elevation"

br_elevation_df <- br_elevation_df %>% 
  filter(x <= -30, x >= -80, y <= 10, y >= -40) %>% 
  mutate(elevation = ifelse(elevation > 3000, 3000, elevation), 
         elevation = ifelse(elevation < 0, NA, elevation))

```
Note: cutting off elevation at 3000m because the highest point in Brazil is this mountain, so as not to get thrown off by the Andes to the west
https://en.wikipedia.org/wiki/Pico_da_Neblina#:~:text=Pico%20da%20Neblina%20(Portuguese%20pronunciation,on%20the%20Brazil%2DVenezuela%20border.



Mapping the elevation
```{r}
ggplot(map) +
  geom_sf(colour = "black") +
  coord_sf(xlim = brazil_window_sf[, "X"],
           ylim = brazil_window_sf[, "Y"]) +
  geom_tile(data = br_elevation_df, 
  			aes(x = x, y = y, fill = elevation)) +
  scale_fill_gradient(low = "white", high = "navy")
```
Save elevation data as a csv
```{r}
write_csv(br_elevation_df, "../DATA/elevation_res5.csv")
```

Now making the boundary box smaller
```{r}
rio_window <- st_sfc(st_point(c(-47.5, -24)), #left, bottom
                        st_point(c(-42.5, -19)), #right, top
                        crs = 4326)

rio_window_sf <- rio_window %>% 
  st_transform(crs = target_crs) %>% 
  st_coordinates()

rio_elev_df <- br_elevation_df %>% 
  filter(y > -24 & y < -19 & x < -42.5 & x > -47.5)
```
Displaying
```{r}
plot_elev <- ggplot(map) +
  geom_tile(data = rio_elev_df, 
  			aes(x = x, y = y, fill = elevation)) +
  scale_fill_gradient(low = "white", high = "navy") +
  geom_sf(alpha = 0.1) +
  coord_sf(xlim = rio_window_sf[, "X"],
           ylim = rio_window_sf[, "Y"]) +
  labs(x = "Longitude", y = "Latitude", fill = "Elevation (m)") +
  theme_bw()
plot_elev
ggsave("~/Documents/extra-projects/un_datathon_2023/VIZ/elevation_plot.jpg")
```

